Variational Monte Carlo on bosonic systems

### 2018-03-25 project_1/article/article.tex
Remove overfull hbox in equation and picture
Use general variational parameter gradient

        From $\mathbf{X}$ we compute the variance and sample covariance between the series itself and a
        one-point lag of the series. This covariance is usually referred to as the first order autocovariance.
        \begin{equation}
            \gamma_i = \frac{1}{n}
                \left(\sum_{k=1}^n\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right)
                \left(\sum_{k=0}^{n-1}\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right).
        \end{equation}

@@

        % Flow chart for description of automatic blocking algorithm
        \begin{figure}
            \centering

                % Defining blocks of flowchart
                \tikzstyle{decision} = [diamond, draw, fill=blue!20,
                    text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
            \tikzstyle{block} = [rectangle, draw, fill=blue!20,
                    text width=11em, text centered, rounded corners, minimum height=4em]
            \tikzstyle{line} = [draw, -latex']

@@

                    % Nodes
                    \node [block] (init) {Start with $\mathbf{X}_i$, $i=0$};
                    \node [block, below of=init] (variance) {Compute variance and autocovariance of $\mathbf{X}_i$, $s_i$ and $\gamma_i$};
                    \node [block, below of=variance] (transform)
                        {Transform data $(\mathbf{X}_{i+1})_k = \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right)$};
                    \node [block, right=1.5cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
                    \node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? };
                    \node [block, below of=x_size] (m_test) {Compute test statistic $M_j$ from $s_j$ and $\gamma_j$ for all $j$.};
                    \node [block, below of=m_test] (magic_numbers) {Find first $k$ so that $M_k \leq q_{d-k}(1-\alpha)$};
                    \node [block, below of=magic_numbers] (return) {Return $s_k^2/n_k$};

@@

        Scanning the entire space of variational parameters can be a somewhat despairing task. To aid us in the search we
        therefore implement the method of gradient descent\footnote{Also known as steepest descent, easily confused
        with the steepest descent method for integation.}. In any optimum we generally have that the derivative should equal
        zero. In our case, we want to find the minimum of the expected local energy with respect to some variational
        parameter $\alpha$, so that
        \begin{equation}
            \nabla_\alpha \expval{E_L(\mathbf{R}; \alpha)} = 0.
        \end{equation}
        The foundation for  the method of gradient descent lies in the fact that some analytic
        \footnote{We use the term analytic quite recklessly.} function $F(\alpha)$ decreases fastest in the direction
        of $-\nabla_\alpha F(\alpha)$. This means that we could eventually get to an ensuing minimum by trailing along the
        pathway of the following recursive relation,
        \begin{equation}
            \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha k} F(\alpha_k).
        \end{equation}
        If $\gamma$ is small enough we should have $f(\alpha_k) \geq F(\alpha_{k+1})$ for all $k$ and hopefully we will
        experience convergence to the desired minimum.

        For our special case we only have one variational parameter, $\alpha$.  The expression for the derivative is,
        \begin{equation}
            \frac{d E_L}{d\alpha} =
                2\left( \expval{\frac{1}{\psi_T}\frac{d\psi_T}{d\alpha} E_L} - \expval{\frac{1}{\psi_T} \frac{d\psi_T}{d\alpha}}\expval{E_L} \right).
        \end{equation}

        INSERT DERIVATION OF LOCAL ENERGY??

>>>>

        From $\mathbf{X}$ we compute the variance and sample covariance between the series itself and a
        one-point lag of the series. This covariance is usually referred to as the first order autocovariance.
        \begin{align}
            \gamma_i
            &= \frac{1}{n}
            \left(
                \sum_{k=1}^n
                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
            \right)
            \nonumber \\
            &\qquad
            \times
            \left(
                \sum_{k=0}^{n-1}
                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
            \right).
        \end{align}

@@

        % Flow chart for description of automatic blocking algorithm
        \begin{figure}
            \centering
            % Defining blocks of flowchart
            \tikzstyle{decision} = [diamond, draw, fill=blue!20,
                text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
            \tikzstyle{block} = [rectangle, draw, fill=blue!20,
                    text width=11em, text centered, rounded corners, minimum height=4em]
            \tikzstyle{line} = [draw, -latex']

@@

                % Nodes
                \node [block] (init) {Start with $\mathbf{X}_i$, $i=0$};
                \node [block, below of=init] (variance) {Compute variance and autocovariance of $\mathbf{X}_i$, $s_i$ and $\gamma_i$};
                \node [block, below of=variance] (transform)
                    {Transform data $(\mathbf{X}_{i+1})_k = \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right)$};
                \node [block, right=1.0cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
                \node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? };
                \node [block, below of=x_size] (m_test) {Compute test statistic $M_j$ from $s_j$ and $\gamma_j$ for all $j$.};
                \node [block, below of=m_test] (magic_numbers) {Find first $k$ so that $M_k \leq q_{d-k}(1-\alpha)$};
                \node [block, below of=magic_numbers] (return) {Return $s_k^2/n_k$};

@@

            Scanning the entire space of variational parameters can be a
            somewhat despairing task. To aid us in the search we therefore
            implement the method of gradient descent\footnote{Also known as
            steepest descent, easily confused with the steepest descent method
            for integation.}. In any optimum we generally have that the
            derivative should equal zero. In our case, we want to find the
            minimum of the expected local energy with respect to some
            variational parameter $\alpha$, so that
            \begin{equation}
                \nabla_\alpha \expval{E_L(\mathbf{R}; \alpha)} = 0.
            \end{equation}
            The foundation for  the method of gradient descent lies in the fact
            that some analytic \footnote{We use the term analytic quite
            recklessly.} function $F(\alpha)$ decreases fastest in the direction
            of $-\nabla_\alpha F(\alpha)$. This means that we could eventually
            get to an ensuing minimum by trailing along the pathway of the
            following recursive relation,
            \begin{equation}
                \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha} F(\alpha_k).
            \end{equation}
            If $\gamma$ is small enough we should have $f(\alpha_k) \geq
            F(\alpha_{k+1})$ for all $k$ and hopefully we will experience
            convergence to the desired minimum.

            For our special case we only have one variational parameter,
            $\alpha$.  The expression for the derivative is,
            \begin{equation}
                \nabla_{\alpha}\expval{E_L}
                =
                2\left(
                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}\expval{E_L}
                \right).
                \label{eq:variational_parameter_gradient}
            \end{equation}

            %TODO: INSERT DERIVATION OF LOCAL ENERGY??
